Model Checking Differentially Private Properties

### 2018-07-27 rebuttal-aplas.txt

First of all, we would like to thank all reviewers' efforts and
comments.

The main contributions of this work are the formalization of privacy
protection mechanisms as Markov chains or Markov Decision Processes,
and its automated analyses. We choose to formalize interactive
mechanisms by non-determinism in Markov Decision Processes and
illustrate our formalization by the above threshold mechanism
(Algorithm 3) for clarity. In the above threshold mechanism, the
incoming queries f_i's are selected dynamically. Later queries may
depend on the results of previous queries. Hence the interaction is
implicit in the choices of incoming queries. It would be great if we
could give our formalization of more explicit interactions such as
those in streaming data. Due to space limit, we had to present a
simpler mechanism with implicit interactions. We also agree a thorough
investigation on the expressiveness of our formalization will be very
useful. It will definitely be an interesting future work. Our
(un)decidability results are for dpCTL*. As long as a mechanism can be
modeled by Markov chains, dpCTL* properties can be verified
automatically. Model checking dpCTL* over Markov Decision Processes on
the other hand is undecidable. Hence interactive mechanisms in general
cannot be verified with our current formalization.

We are certainly aware of various language-based techniques for
verifying differential privacy. Just like classical
programs/properties, we see both techniques are complementary.
Classical model checking aims to reduce human efforts by focusing on
abstract models with automated verification. It has helped programmers
verify simpler programs and properties reasonably successfully over
decades of research efforts. Our work has the same position. It does
not aim to verify all possible differentially private programs.
Instead, it hopefully will solve simpler problems so that programmers
can have more time for language-based techniques.

### 2018-07-28 rebuttal-aplas.txt

First of all, we would like to thank all reviewers' efforts and
comments.

The main contributions of this work are the formalization of privacy
protection mechanisms as Markov chains or Markov Decision Processes,
and its automated analyses. We choose to formalize interactive
mechanisms by non-determinism in Markov Decision Processes and
illustrate our formalization by the above threshold mechanism
(Algorithm 3) for clarity. In the above threshold mechanism, the
incoming queries f_i's are selected dynamically. Later queries may
depend on the results of previous queries. Hence the interaction is
implicit in the choices of incoming queries. It would be great if we
could give our formalization of more explicit interactions such as
those in streaming data. Due to space limit, we had to present a
simpler mechanism with implicit interactions. We also agree a thorough
investigation on the expressiveness of our formalization will be very
useful. It will definitely be an interesting future work. Our
(un)decidability results are for dpCTL*. As long as a mechanism can be
modeled by Markov chains, dpCTL* properties can be verified
automatically. Model checking dpCTL* over Markov Decision Processes on
the other hand is undecidable. Hence interactive mechanisms in general
cannot be verified with our current formalization.

We are aware of various language-based techniques for verifying
differential privacy (of course). Just like classical 
programs/properties, we see both techniques are complementary.
Classical model checking aims to reduce human efforts by focusing on
abstract models with automated verification. It has helped programmers
verify simpler programs and properties reasonably successfully over
decades of research efforts. Our work has the same position. It does
not aim to verify all possible differentially private programs.
Instead, it hopefully will solve simpler problems so that programmers
can have more time for language-based techniques.

F\phi (finally \phi) and G\phi (globally \phi) are two common temporal
properties used in PCTL*. They are syntactic sugars (page 8).
The time complexity in Proposition 1 is polynomial in
|S|, which includes |S|^2. Actually the computations in PCTL* alone
would take more than |S|^2. Since differential privacy requires the
indistinguishable probabilities on all neighbors, thus we need to
compute the probabilities and compare between neighbors. The
relational assertion just tells us which pairs should be neighbors, without
guaranteeing the probabilities. The other two examples actually show
us explicit models and computations, though, which are much easier and
obvious. There exist incoming edges for t1’r2’, such as from t1’r0, t1’r1
and so on. We explain these transitions in section 7.1 and leave out
some edges for clarity.

### 2018-07-28 above-threshold.tex

Now, the non-deterministic choices in the two computation cannot be 
independent. Recall that the sensitivity of each query is
$1$. If the top computation moves to the state, say, $t'_1r_0$, it
means the next query result on the dataset is $0$. Subsequently, the
bottom computation can only move to $\underline{t}'_j\underline{r}_0$
or $\underline{t}'_j\underline{r}_1$ depending on its perturbed
threshold. This is where actions are useful. Define the actions
$\{ m\underline{n} : |m - n| \leq 1 \}$. The action
$m\underline{n}$ represents that the next query result for the dataset
and its neighbor are $m$ and $n$ respectively. For instance, the
non-deterministic choice from $t'_1r'_0$ to $t'_1r_0$ is associated
with two actions $0\underline{0}$ and $0\underline{1}$ (but not
$0\underline{2}$). Similarly, the choice from
$\underline{t}'_2\underline{r}'_1$ to
$\underline{t}'_2\underline{r}_0$ is associated with the actions
$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). 
Assume the perturbed thresholds of the top and bottom computation are
$i$ and $j$ respectively. On the action $0\underline{0}$, the top
computation moves to $t'_ir_0$ and 
the bottom computation moves to $\underline{t}'_j\underline{r}_0$ 
Actions make sure the two computation of neighbors
is modeled properly. Now consider the action sequence
$-,-,0\underline{1},-,2\underline{2},-,2\underline{1}$ from the states
$t_0r_1$ and $\underline{t}_0\underline{r}_0$ (``$-$'' represents the
purely probabilistic action). Together with the first query results, 
it denotes four consecutive query results $1$, $0$, $2$, $2$ on the
top computation, and $0$, $1$, $2$, $1$ on the bottom computation.
Each action sequence models two sequences of query results: one on the
top, the other on the bottom computation. Moreover, the difference of
the corresponding query results on the two computation is at most one
by the definition of the action set. Any sequence of adaptive query
results is hence formalized by an action sequence in our model.

>>>>

Now, the non-deterministic choices in the two computation cannot be 
independent. Recall that the sensitivity of each query is
$1$. If the top computation moves to the state, say, $t'_1r_0$, it
means the next query result on the dataset is $0$. Subsequently, the
bottom computation can only move to $\underline{t}'_j\underline{r}_0$
or $\underline{t}'_j\underline{r}_1$ depending on its perturbed
threshold. This is where actions are useful. Define the actions
$\{ m\underline{n} : |m - n| \leq 1 \}$. The action
$m\underline{n}$ represents that the next query result for the dataset
and its neighbor are $m$ and $n$ respectively. For instance, the
non-deterministic choice from $t'_1r'_0$ to $t'_1r_0$ is associated
with two actions $0\underline{0}$ and $0\underline{1}$ (but not
$0\underline{2}$). Similarly, the choice from
$\underline{t}'_2\underline{r}'_1$ to
$\underline{t}'_2\underline{r}_0$ is associated with the actions
$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). 
Assume the perturbed thresholds of the top and bottom computation are
$i$ and $j$ respectively. On the action $0\underline{0}$, the top
computation moves to $t'_ir_0$ and 
the bottom computation moves to $\underline{t}'_j\underline{r}_0$.
Actions make sure the two computation of neighbors
is modeled properly. Now consider the action sequence
$-,-,0\underline{1},-,2\underline{2},-,2\underline{1}$ from the states
$t_0r_1$ and $\underline{t}_0\underline{r}_0$ (``$-$'' represents the
purely probabilistic action). Together with the first query results, 
it denotes four consecutive query results $1$, $0$, $2$, $2$ on the
top computation, and $0$, $1$, $2$, $1$ on the bottom computation.
Each action sequence models two sequences of query results: one on the
top, the other on the bottom computation. Moreover, the difference of
the corresponding query results on the two computation is at most one
by the definition of the action set. Any sequence of adaptive query
results is hence formalized by an action sequence in our model.
